NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration

### 2022-08-31 main.tex
Updates from Overleaf

%averaging Here we demonstrate the power of the NMRlipids databank by analysing how 
Water permeabilities calculated from all the simulations in the NMRlipids databank as a function of area per lipids, thickness, and acyl chain order are shown in  Fig.~\ref{fig:permeability} (a-c). To illustrate how permeability depends these properties, average over systems with fixed values of area per lipids, thickness, and acyl chain order are also shown in the figure. Permeability increases with increasing area per lipid and decreases with increasing membrane thickness and acyl chain order. Previous simulation and experimental studies have reported linear dependence of permeability on both area per lipid and thickness for limited set of lipid mixtures but not always for all types of lipid mixtures \cite{mathai08,frallicciardi22}. Our results suggest stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$, but the dependence may appear linear above and below these values as shown in the insets of Figs. \label{fig:permeability} a) and b). As expected, the permeability increases with increasing temperature in Figs. \label{fig:permeability} d). The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation. The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=180mm]{Figures/permeation2.pdf}
    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
      a) area per lipid,
      b) thickness,
      and c) temperature. 
      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
    \label{fig:permeability}
\end{figure}

To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane. The results at different hydration levels are shown in Fig.~\ref{fig:permeability} e) together with the experimental data~\cite{??}. The experimental water diffusion increases with increasing hydration level toward the value for bulk water (?? in ??K). Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion. However, simulation results are closer to experiments with low hydration levels. To estimate the diffusion anistoropy, we translated the water permation to diffusion coefficent through multilamellar stack using the Tanner equation~\cite{??}. The results in

>>>>

%averaging Here we demonstrate the power of the NMRlipids databank by analysing how 
Water permeabilities calculated from all the simulations in the NMRlipids databank as a function of area per lipids, thickness, and acyl chain order are shown in  Fig.~\ref{fig:permeability} (a-c). To illustrate how permeability depends these properties, average over systems with fixed values of area per lipids, thickness, and acyl chain order are also shown in the figure. Permeability increases with increasing area per lipid and decreases with increasing membrane thickness and acyl chain order. Previous simulation and experimental studies have reported linear dependence of permeability on both area per lipid and thickness for limited set of lipid mixtures but not always for all types of lipid mixtures \cite{mathai08,frallicciardi22}. Our results suggest stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$, but the dependence may appear linear above and below these values as shown in the insets of Figs. \label{fig:permeability} a) and b). As expected, the permeability increases with increasing temperature in Figs. \label{fig:permeability} d). The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation. The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}. 
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=180mm]{Figures/permeation2.pdf}
    \caption{(b-c) Water permeation through lipid membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order. 
      Inserts in c) and d) show the region where the dependence could be considered approximately linear.
      Insert in a) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.
      (f-g) Diffusion anistoropy of water as a function of thickness and area per lipid.
      (h) Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.}
    \label{fig:permeability}
\end{figure}

To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane. The results at different hydration levels are shown in Fig.~\ref{fig:permeability} h) together with the experimental data~\cite{rudakova04}. The experimental water diffusion increases with increasing hydration level toward the value for bulk water (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K)\cite{khakimov08}. Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion. However, simulation results are closer to experiments with low hydration levels. To estimate the diffusion anistoropy, we translated the water permeation to diffusion coefficient through multilamellar stack using the Tanner equation~\cite{??}. The resulting diffusion coefficients trough the membranes are approximately five orders of magnitude slower than along the membrane which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}. Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. Significant dependence of diffusion anisotropy on membrane thickness and area per lipid are observed in Figs.~\ref{fig:permeability} f) and g). The anisotropy becomes linearly stronger with decreasing area per lipid while the dependency is stronger with increasing thickness. These results can be explained by the increasing and decreasing permeabilities with thickness and are per lipid (Figs \ref{fig:permeability} c) and d), respectively, while lateral diffusion remains approximately constant (Fig. \ref{??} a) and c)).
